---
title: "Exploratory analysis"
author: "Simon"
date: "2024-11-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(readr)
library(readxl)
library(dplyr)
library(skimr)
library(binsreg)
library(splines)
library(sf)
library(estimatr)
library(fixest)
```


```{r}
setwd("G:/My Drive/In Course Project/")
```

Immigration Data (Inflow Data)
2019/2020
```{r}
data2=read_csv("G:/My Drive/In Course Project/countyinflow1920.csv")
inflow_1920 = data2[grep("Total Migration-US and Foreign",data2$y1_countyname), ]
inflow_1920 = inflow_1920 |>
  mutate(FIPS=paste(y2_statefips,y2_countyfips,sep = ""))
inflow_1920 = inflow_1920|>
  filter(y2_countyfips !="000")
# Removes the Total Migration-US and Foreign 
inflow_1920=inflow_1920|>
  mutate(y1_countyname=str_replace(y1_countyname,"Total Migration-US and Foreign",""))
colnames(inflow_1920) = c("Y2_STATEFIPS","Y2_COUNTYFIPS","Y1_STATEFIPS","Y1_COUNTYFIPS","Y1_STATE","Y1_COUNTYNAME","NO_OF_RETURN(INFLOW)","NO_OF_INDIVIDUALS(INFLOW)","AGI_INFLOW","FIPS" )
```
2018/2019
```{r}
data3=read_csv("G:/My Drive/In Course Project/countyinflow1819.csv")
inflow_1819 = data3[grep("Total Migration-US and Foreign",data3$y1_countyname), ]
inflow_1819 = inflow_1819 |>
  mutate(FIPS=paste(y2_statefips,y2_countyfips,sep = ""))
inflow_1819 = inflow_1819|>
  filter(y2_countyfips !="000")
# Removes the Total Migration-US and Foreign 
inflow_1819=inflow_1819|>
  mutate(y1_countyname=str_replace(y1_countyname,"Total Migration-US and Foreign",""))
colnames(inflow_1819) =  c("Y2_STATEFIPS","Y2_COUNTYFIPS","Y1_STATEFIPS","Y1_COUNTYFIPS","Y1_STATE","Y1_COUNTYNAME","NO_OF_RETURN(INFLOW)","NO_OF_INDIVIDUALS(INFLOW)","AGI_INFLOW","FIPS" )

```
2017/2018
```{r}
data4=read_csv("G:/My Drive/In Course Project/countyinflow1718.csv")
inflow_1718 = data4[grep("Total Migration-US and Foreign",data4$y1_countyname), ]
inflow_1718 = inflow_1718 |>
  mutate(FIPS=paste(y2_statefips,y2_countyfips,sep = ""))
inflow_1718 = inflow_1718|>
  filter(y2_countyfips !="000")
# Removes the Total Migration-US and Foreign 
inflow_1718=inflow_1718|>
  mutate(y1_countyname=str_replace(y1_countyname,"Total Migration-US and Foreign",""))
colnames(inflow_1718) =  c("Y2_STATEFIPS","Y2_COUNTYFIPS","Y1_STATEFIPS","Y1_COUNTYFIPS","Y1_STATE","Y1_COUNTYNAME","NO_OF_RETURN(INFLOW)","NO_OF_INDIVIDUALS(INFLOW)","AGI_INFLOW","FIPS" )

```
2016/2017
```{r}
data5=read_csv("G:/My Drive/In Course Project/countyinflow1617.csv")
inflow_1617 = data5[grep("Total Migration-US and Foreign",data5$y1_countyname), ]
inflow_1617 = inflow_1617 |>
  mutate(FIPS=paste(y2_statefips,y2_countyfips,sep = ""))
inflow_1617 = inflow_1617|>
  filter(y2_countyfips !="000")
# Removes the Total Migration-US and Foreign 
inflow_1617=inflow_1617|>
  mutate(y1_countyname=str_replace(y1_countyname,"Total Migration-US and Foreign",""))
colnames(inflow_1617) =  c("Y2_STATEFIPS","Y2_COUNTYFIPS","Y1_STATEFIPS","Y1_COUNTYFIPS","Y1_STATE","Y1_COUNTYNAME","NO_OF_RETURN(INFLOW)","NO_OF_INDIVIDUALS(INFLOW)","AGI_INFLOW","FIPS" )

```
2015/2016
```{r}
data6=read_csv("G:/My Drive/In Course Project/countyinflow1516.csv")
inflow_1516 = data6[grep("Total Migration-US and Foreign",data6$y1_countyname), ]
inflow_1516 = inflow_1516 |>
  mutate(FIPS=paste(y2_statefips,y2_countyfips,sep = ""))
inflow_1516 = inflow_1516|>
  filter(y2_countyfips !="000")
# Removes the Total Migration-US and Foreign 
inflow_1516=inflow_1516|>
  mutate(y1_countyname=str_replace(y1_countyname,"Total Migration-US and Foreign",""))
colnames(inflow_1516) =  c("Y2_STATEFIPS","Y2_COUNTYFIPS","Y1_STATEFIPS","Y1_COUNTYFIPS","Y1_STATE","Y1_COUNTYNAME","NO_OF_RETURN(INFLOW)","NO_OF_INDIVIDUALS(INFLOW)","AGI_INFLOW","FIPS" )

```
2014/2015
```{r}
data7=read_csv("G:/My Drive/In Course Project/countyinflow1415.csv")
inflow_1415 = data7[grep("Total Migration-US and Foreign",data7$y1_countyname), ]
inflow_1415 = inflow_1415 |>
  mutate(FIPS=paste(y2_statefips,y2_countyfips,sep = ""))
inflow_1415 = inflow_1415|>
  filter(y2_countyfips !="000")
# Removes the Total Migration-US and Foreign 
inflow_1415=inflow_1415|>
  mutate(y1_countyname=str_replace(y1_countyname,"Total Migration-US and Foreign",""))
colnames(inflow_1415) =  c("Y2_STATEFIPS","Y2_COUNTYFIPS","Y1_STATEFIPS","Y1_COUNTYFIPS","Y1_STATE","Y1_COUNTYNAME","NO_OF_RETURN(INFLOW)","NO_OF_INDIVIDUALS(INFLOW)","AGI_INFLOW","FIPS" )

```
2013/2014
```{r}
data8=read_csv("G:/My Drive/In Course Project/countyinflow1314.csv")
inflow_1314 = data8[grep("Total Migration-US and Foreign",data8$y1_countyname), ]
inflow_1314 = inflow_1314 |>
  mutate(FIPS=paste(y2_statefips,y2_countyfips,sep = ""))
inflow_1314 = inflow_1314|>
  filter(y2_countyfips !="000")
# Removes the Total Migration-US and Foreign 
inflow_1314=inflow_1314|>
  mutate(y1_countyname=str_replace(y1_countyname,"Total Migration-US and Foreign",""))
colnames(inflow_1314) =  c("Y2_STATEFIPS","Y2_COUNTYFIPS","Y1_STATEFIPS","Y1_COUNTYFIPS","Y1_STATE","Y1_COUNTYNAME","NO_OF_RETURN(INFLOW)","NO_OF_INDIVIDUALS(INFLOW)","AGI_INFLOW","FIPS" )

```
2012/2013
```{r}
data9=read_csv("G:/My Drive/In Course Project/countyinflow1213.csv")
inflow_1213 = data9[grep("Total Migration-US and Foreign",data9$y1_countyname), ]
inflow_1213 = inflow_1213 |>
  mutate(FIPS=paste(y2_statefips,y2_countyfips,sep = ""))
inflow_1213 = inflow_1213|>
  filter(y2_countyfips !="000")
# Removes the Total Migration-US and Foreign 
inflow_1213=inflow_1213|>
  mutate(y1_countyname=str_replace(y1_countyname,"Total Migration-US and Foreign",""))
colnames(inflow_1213) = c("Y2_STATEFIPS","Y2_COUNTYFIPS","Y1_STATEFIPS","Y1_COUNTYFIPS","Y1_STATE","Y1_COUNTYNAME","NO_OF_RETURN(INFLOW)","NO_OF_INDIVIDUALS(INFLOW)","AGI_INFLOW","FIPS" )

```
housing data
```{r}
new_data_1=read_csv("G:/My Drive/In Course Project/UADAggs_county_v3_2.csv")

filtered_data <- new_data_1 |>
#keeps only observation of the fifth quarter and purchase and drops the 2023 observations.  
  filter(QUARTER == 5, PURPOSE == "Purchase",YEAR != 2023)|>
  arrange(YEAR,GEONAME)|> # 
  filter(SERIES=="Count of Appraisals")|> #Keeps the observations with series ID equals count of appraisal
  arrange(STATEFIPS)
#Removes row 31135 to row 31917 from the data set
dropped_data = filtered_data[-c(31135:31917), ]
# Drops the columns 
housing_prices=dropped_data|>
  select(-c("TRACT","FREQUENCY", "METRO","APPRAISALSOURCE","SERIES","SERIESID","SOURCE","CHARACTERISTIC1", "GEOLEVEL", "CATEGORY1"))
colnames(housing_prices) = c("Y1_COUNTYNAME","STATEPOSTAL","STATEFIPS","FIPS","PURPOSE","YEAR","QUARTER","SUPPRESSED","VALUE")


```
```{r}
d_13 = housing_prices|>
  filter(YEAR==2013)
d_14= housing_prices|>
  filter(YEAR==2014)
d_15 = housing_prices|>
  filter(YEAR==2015)
d_16 = housing_prices|>
  filter(YEAR==2016)
d_17 = housing_prices|>
  filter(YEAR==2017)
d_18 = housing_prices|>
  filter(YEAR==2018)
d_19 = housing_prices|>
  filter(YEAR==2019)
d_20 = housing_prices|>
  filter(YEAR==2020)


```

Merging the data sets in bits
```{r}
# 1st
d_13 = housing_prices|>
  filter(YEAR==2013)
merge_1 = inflow_1213|>
  left_join(d_13, by= "FIPS")

# 2nd
d_14= housing_prices|>
  filter(YEAR==2014)
merge_2 = inflow_1314|>
  left_join(d_14, by = "FIPS")

#3rd 
d_15 = housing_prices|>
  filter(YEAR==2015)
merge_3 = inflow_1415|>
  left_join(d_15, by = "FIPS")

#4th
d_16 = housing_prices|>
  filter(YEAR==2016)
merge_4 = inflow_1516|>
  left_join(d_16, by = "FIPS")

#5th 
d_17 = housing_prices|>
  filter(YEAR==2017)
merge_5 = inflow_1617|>
  left_join(d_17, by = "FIPS")

#6th 
d_18 = housing_prices|>
  filter(YEAR==2018)
merge_6 =inflow_1718|>
  left_join(d_18, by = "FIPS")

#7th 
d_19 = housing_prices|>
  filter(YEAR==2019)
merge_7 = inflow_1819|>
  left_join(d_19, by = "FIPS")

#8th 
d_20 = housing_prices|>
  filter(YEAR==2020)
merge_8 = inflow_1920|>
  left_join(d_20, by = "FIPS")

```

Combining the dataset using rbind
```{r}
main_data = rbind(merge_1,merge_2,merge_3,merge_4,merge_5,merge_6,merge_7,merge_8)

```

some exploratory analysis
```{r}
summary(main_data)
```
```{r}
skim(main_data)
```

removing the NAs(missing) from the value variable
```{r}
clean_data = main_data[!is.na(main_data$VALUE),]
clean_data$LOG_VALUE = log(clean_data$VALUE + 1)
clean_data$LOG_MIGRATION = log(clean_data$`NO_OF_INDIVIDUALS(INFLOW)` + 1)
```



```{r}
ggplot(clean_data,aes(x=VALUE))+ 
  geom_histogram(boundary=0,bindwith = 2)
```
This shows the distribution of value(housing prices) variable, which highly right skewed. Most of the data are concentrated on the left side of the distribution. This points out that significant portion of the values in the dataset are small with few proportion been large. 

```{r}
ggplot(clean_data,aes(x=`NO_OF_RETURN(INFLOW)`))+ 
  geom_histogram(boundary=0,bindwith = 2)
```

```{r}
ggplot(clean_data,aes(x=`NO_OF_INDIVIDUALS(INFLOW)`))+ 
  geom_histogram(boundary=0,bindwith = 2)
```
This shows the distribution of number of individual(inflow)i.e. migrants variable, which highly right skewed. Most of the data are concentrated on the left side of the distribution. 



Applying log transformation to variables because they are right skewed.Applying log will reduce the skewness, compressing the larger values and spreading out the smaller ones. This will better reflect the central tendency.
```{r}
ggplot(clean_data,aes(x=LOG_VALUE))+ 
  geom_histogram(boundary=0,bindwith = 0.5)
```

```{r}
ggplot(clean_data,aes(x=LOG_MIGRATION))+ 
  geom_histogram(boundary=0,bindwith = 0.5)
```


```{r}
ggplot(clean_data, aes(x=LOG_MIGRATION, y =LOG_VALUE)) +
  geom_point(alpha = 0.1) 
```

There is a clear positive linear relationship (upward sloping pattern) between number of individuals inflow (migrants) and housing prices. This indicates as the number of migrants increases, housing prices also tends to increase. The points are tightly clustered around a linear trend suggesting a high correlation.


```{r}
binsreg(log(clean_data$VALUE),log(clean_data$`NO_OF_INDIVIDUALS(INFLOW)`))
```
There is an overall positive relationship between immigration and housing prices, as indicated by the upward trend. As the NO_OF_INDIVIDUALS(INFLOW) increases, VALUE also tends to increase, suggesting a direct association(correlation) between these two variables. This also affirms the relationship between the variables by the scatter graph even though this show a non - linear relationship

Using a local regression 
```{r}
# Create a scatter plot with a smooth trend line using ggplot with clean_data data set, mapping 'NO_OF_INDIVIDUAL(INFLOW)' to x-axis and 'VALUE' to y-axis
ggplot(clean_data,aes(x=LOG_MIGRATION,y= LOG_VALUE)) +
  geom_point(alpha=0.3, size=1.5)+
# Add a smooth line using LOESS smoothing with linear fitting  
  geom_smooth(method = "loess")
```
The graph shows a positive relationship (upward trend) between number of individuals (inflow) and housing prices, signifying that when migration(inflow) increases housing prices.The relationship is not a linear one but due to the smoothness some of the variations are been averaged out hence we do not have enough information about some key variations.


creating a bin scatter
```{r}
#  Create 50 bins based on the 'NO_OF_INDIVIDUALS(INFLOW)' variable and calculate binned means
clean_data =clean_data|>
  mutate(INDIVIDUAL_bins=factor(ntile(LOG_MIGRATION, 50)))

# Group the data by the created bins and calculate the average 'NO_OF_INDIVIDUALS' and 'VALUE' within each bin
binned_data = clean_data |>
  group_by(INDIVIDUAL_bins) |>
  # Calculate the mean 'VALUE' for each bin, ignoring missing values
  mutate(VALUE_binned = mean(LOG_VALUE,na.rm = T),
# Calculate the mean 'NO_OF_INDIVIDUALS' for each bin, ignoring missing values
         INDIVIDUALS_binned = mean(LOG_MIGRATION, na.rm = T))

 
```

Creating the graph
```{r}
# Create a scatter plot with the clean_data and binned data using 
ggplot
ggplot(binned_data) +
  geom_point(aes(x = LOG_MIGRATION, y = LOG_VALUE, color = INDIVIDUAL_bins), alpha = 0.3) +
  geom_point(aes(x = INDIVIDUALS_binned, y = VALUE_binned), size = 3, color = "navy") +  # Binned Averages
  labs(x = "Number of People inflow", y = "Average House Prices") +
  theme_minimal()
```
The bin scatter plot suggests a positive correlation between the number of people inflow and average house prices. As the number of people inflow increases, average house prices generally tend to increase as well. This pattern suggests that higher inflows of people are associated with higher house prices, possibly due to demand pressures in areas with higher population inflow. 

creating a local regression using the bin
```{r}
# Create a smooth trend line plot using ggplot2
ggplot(binned_data,aes(x=LOG_MIGRATION, y= LOG_VALUE))+
  geom_smooth()# Add a smooth trend line to visualize the overall relationship between NO_OF_INDIVIDUALS and VALUE
```
The CEF shows a clear linear upward sloping trend(positive relationship between migrations and housing prices). This suggests that higher levels of migrations are generally associated with higher house value.



Trends of individual migration (Number of individuals inflow)
```{r}
ggplot(clean_data, aes(x = YEAR, y = LOG_MIGRATION, group = Y1_COUNTYNAME.x, color = Y1_COUNTYNAME.x)) +
  geom_point() +
  labs(title = "Trend of NO_OF_INDIVIDUALS(INFLOW) by Year", x = "YEAR", y = "NO_OF_INDIVIDUALS(INFLOW)") +
  theme(legend.position = "none")
```

Plotting time series trend for the data 
```{r}
ggplot(clean_data, aes(x = YEAR, y = LOG_VALUE, group = Y1_COUNTYNAME.x, color = Y1_COUNTYNAME.x)) +
  geom_point() +
  labs(title = "Trend of VALUE by Year", x = "YEAR", y = "VALUE") +
  theme(legend.position = "none")
```



Box Plot for counties with the highest VALUE
```{r}
top_counties = clean_data |>
  group_by(Y1_COUNTYNAME.x)|>
  summarise(mean_value = mean(LOG_VALUE, na.rm = T))|>
  arrange(desc(mean_value))|>
  top_n(10,mean_value)

# Filters the original data to only include top counties
filtered_counties= clean_data|>
  filter(Y1_COUNTYNAME.x %in% top_counties$Y1_COUNTYNAME.x)

# Box plot for top counties 
ggplot(filtered_counties, aes(x = reorder(Y1_COUNTYNAME.x, -LOG_VALUE), y = LOG_VALUE)) +
  geom_boxplot(fill = "lightgreen") +
  labs(title = "Box Plot of VALUE for Top 10 Counties", x = "County", y = "VALUE") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
The boxes show the interquartile range (IQR), which is the range from the 25th percentile (bottom edge of the box) to the 75th percentile (top edge of the box).The line within each box represents the median VALUE for that county. Whiskers extend from the box to indicate the range of values within 1.5 times the IQR. Points outside this range are potential outliers and are shown as individual dots.
Maricopa County has the highest median and overall spread for VALUE
Harris County has a wide IQR and median value but shows a larger spread than most other counties, suggesting a significant variation in VALUE.
San Diego County also shows a wide range, with both a high median and notable spread in VALUE.
Counties like King County, Tarrant County, Collin County, Wake County, and San Bernardino County have relatively narrow distributions, with lower median values and smaller ranges, indicating less variability in VALUE.
Counties like King County, Tarrant County, Collin County, Wake County, and San Bernardino County have relatively narrow distributions, with lower median values and smaller ranges, indicating less variability in VALUE.
Outliers are present in Collin County and Wake County, represented by individual dots outside the whiskers. These are VALUE observations that are higher than typical values for these counties.
This plot indicates that VALUE(housing price) varies significantly across these counties. Counties like Maricopa and San Diego have high values and wide ranges, while others (like Collin, Wake, and San Bernardino) have lower values and less variation


Box Plot for counties with the highest of individual inflow  
```{r}
top2_counties = clean_data |>
  group_by(Y1_COUNTYNAME.x)|>
  summarise(mean_value = mean(LOG_MIGRATION, na.rm = T))|>
  arrange(desc(mean_value))|>
  top_n(10,mean_value)

# Filters the original data to only include top counties
filter_counties= clean_data|>
  filter(Y1_COUNTYNAME.x %in% top2_counties$Y1_COUNTYNAME.x)

# Box plot for top counties 
ggplot(filter_counties, aes(x = reorder(Y1_COUNTYNAME.x, -LOG_MIGRATION), y = LOG_MIGRATION)) +
  geom_boxplot(fill = "lightgreen") +
  labs(title = "Box Plot of Individual Inflow for Top 10 Counties", x = "County", y = "Number of Individual Inflow") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```



Mean Data by state and plot
```{r}
#Group the data by state and calculates the mean of VALUE
state_summary = clean_data|>
  group_by(STATEPOSTAL)|>
  summarise(mean_value = mean(LOG_VALUE,na.rm = T))

#Plot the box plot
ggplot(state_summary, aes(x = reorder(STATEPOSTAL, -mean_value), y= mean_value))+
   geom_point(fill = "lightblue", size = 3) +
  labs(title = "Scatter Plot of Average VALUE by State", x = "State", y = "VALUE") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


By the State level
```{r}
# Calculate summary statistics for mean VALUE and mean inflow
state_summary2 = clean_data |>
  group_by(STATEPOSTAL) |>
  summarise(
    mean_value = mean(LOG_VALUE, na.rm = TRUE),
    mean_inflow = mean(LOG_MIGRATION, na.rm = TRUE)
  )
```

Taking the top 10 states for both Value and number of individuals inflow
```{r}
# Filter for the top 10 states by mean VALUE and mean inflow
top_value_states = state_summary2 |>
  arrange(desc(mean_value)) |>
  slice(1:10)

top_inflow_states = state_summary2 |>
  arrange(desc(mean_inflow)) |>
  slice(1:10)
```


Graph 
```{r}
# Plot for top 10 states by mean inflow
s1 = clean_data|>
  filter(STATEPOSTAL %in% top_inflow_states$STATEPOSTAL)

ggplot(s1, aes(x = reorder(STATEPOSTAL, -LOG_MIGRATION), y = LOG_MIGRATION)) +
  geom_boxplot(fill = "lightgreen") +
  labs(title = "Top 10 States by Inflow Distribution", x = "State", y = "MIGRATION") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Plot the top ten states based of mean of VALUE
```{r}
s2 = clean_data|>
  filter(STATEPOSTAL %in% top_value_states$STATEPOSTAL)

ggplot(s2, aes(x = reorder(STATEPOSTAL, -LOG_VALUE), y = LOG_VALUE)) +
  geom_boxplot(fill = "lightgreen") +
  labs(title = "Top 10 States by VALUE Distribution", x = "State", y = "VALUE") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Using a local regression 
```{r}
ggplot(s1, aes(x = reorder(STATEPOSTAL, -LOG_VALUE), y = LOG_VALUE)) +
  geom_point(alpha = 0.2) +
  geom_smooth()+  # Local regression line
  labs(title = "Top 10 States by VALUE with Local Regression", x = "State", y = "VALUE") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
ggplot(s2, aes(x = reorder(STATEPOSTAL, -LOG_VALUE), y = LOG_VALUE)) +
  geom_point(alpha = 0.2) +
  geom_smooth()+  # Local regression line
  labs(title = "Top 10 States by VALUE with Local Regression", x = "State", y = "VALUE") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Polynomial Regression 
```{r}
Poly_reg = feols(VALUE ~ poly(`NO_OF_INDIVIDUALS(INFLOW)`, 4), data = clean_data)

summary(Poly_reg)
```
Predicting the fitted values
```{r}
# predict to get fitted values
poly_predict = predict(Poly_reg,data = clean_data,interval = "confidence")
# Combine the original data with the predicted values and confidence intervals
poly_fitted = cbind(clean_data,poly_predict)


# Create a plot to visualize the polynomial regression fit and confidence intervals
ggplot(poly_fitted, aes(x =`NO_OF_INDIVIDUALS(INFLOW)`))+ theme() +
  geom_point(aes(y=VALUE),alpha = 0.1)+  # Scatter plot of actual data 
  # Confidence interval ribbon around the fitted line
  geom_ribbon(aes(ymin = ci_low, ymax = ci_high), fill = "coral") +
  # Fitted line from the polynomial regression
  geom_line(aes(y = fit),size = 1,color = "orange") + 
  labs(x = "Number of Immigrants", y = "Housing Prices", title = "Forth-Order Polynomial Regression") +
  theme_minimal()
```
This graph shows a scatter plot of the relationship between Number of Immigrants (on the x-axis) and Housing Prices (on the y-axis), with a fourth-order polynomial regression line (in orange) fitted to the data. The graph show an overall positive trend between the two variables, where housing prices increase as the number of immigrants grows, but with considerable spread. The polynomial regression captures all the complexities that would have otherwise be smoothen out by fitting a linear regression. At the initial stages the housing prices increase as the number of immigrants rises. The curve begins to flatten at the middle part (showing slow rate of increase in housing prices) and peaks again. And at the tail end has a slight downward slope indicating that increase in immigrant no longer had an impact on housing prices.




Bin Regression
```{r}
# Perform a binned regression using the feols function
binned_reg = feols(LOG_VALUE~as_factor(ntile(LOG_MIGRATION,10)), data = clean_data)
# Display the summary of the regression model
summary(binned_reg)

# Predict values using the binned regression model and add confidence intervals
binned_reg_predict=predict(binned_reg,data=clean_data,interval = "confidence")
# Combine the original data with the predicted values and confidence intervals
binned_reg_fitted= cbind(clean_data,binned_reg_predict)

# Create a plot to visualize the binned regression fit and confidence intervals
ggplot(binned_reg_fitted,aes(x=LOG_MIGRATION))+ theme_light() +
   # Scatter plot of actual data points with low opacity
  geom_point(aes(y = LOG_VALUE), alpha = 0.1) +
   # Confidence interval ribbon around the fitted line
  geom_ribbon(aes(ymin = ci_low, ymax = ci_high), fill = "red") +
  # Fitted line from the binned regression model
  geom_line(aes(y = fit), size = 1, color = "lightgreen")

```

Cubic Splines 
```{r}
# Perform a piece wise cubic regression using B-splines
pw_cubic=feols(VALUE~ splines::bs(`NO_OF_INDIVIDUALS(INFLOW)`,df=5),
               clean_data)
# Display the summary of the regression model
summary(pw_cubic)


# Generate fitted values for cubic splines using predict
pw_cubic_predict = predict(pw_cubic,data=clean_data, interval = "confidence")
pw_cubic_fitted=cbind(clean_data,pw_cubic_predict)

# Create a plot to visualize the piece wise cubic regression fit and confidence intervals
ggplot(pw_cubic_fitted, aes(x = `NO_OF_INDIVIDUALS(INFLOW)`)) + theme_light()+
  # Scatter plot of original data points 
  geom_point(aes(y=VALUE),alpha = 0.1)+ 
  # Confidence interval ribbon around the fitted line
    geom_ribbon(aes(ymin = ci_low, ymax = ci_high), fill = "coral", alpha=2)+
  # Fitted line from the piece wise cubic regression model
  geom_line(aes(y =fit), size = 1,color = "green")+ 
  labs(x = "Number of Immigrants", y = "Housing Prices", title = "Cubic Splines Regression (5 Degrees of Freedom)") +
  theme_minimal()
```

