---
title: "Final Project and Econometric Analysis"
author: "Simon and Steven"
date: "2024-12-05"
output: html_document
---



### Topic

### The impact of Migration on Housing Prices in the US

### Motivation

Migration is the movement of people from one place to another. This can happen internationally when people move from one country to another, or it can happen internally as people move within the same country. Migration can be temporary or permanent, voluntary or forced. It can also be seasonal.

Numerous factors, such as economic opportunities, escaping conflict or persecution, reuniting with family, or educational purposes, inform migration decisions for people moving between nations. Family, employment, educational pursuits, or other interests are some of the underlying reasons people may choose to relocate within the same country. 

Housing prices stand for the cost of purchasing properties, such as houses, apartments, or condominiums. Migration significantly impacts all facets of the US economy, including the housing market. We want to investigate the impact of migrants’ influx (both domestic and foreign) on local housing prices, considering also the demand and supply factors in the market such as interest rates. 


### Why It Matters

Firstly, it is crucial for policymakers as it aids them in making informed decisions regarding migration, which in turn helps to stabilize housing prices, ensuring affordability and a high standard of living. 

Furthermore, it can facilitate the establishment of a stable market, as both homeowners and renters have a reasonable insight into the fluctuations in the housing market. 

Finally, it can also function as a guide for urban planning and infrastructure development.



### Research Questions

1. What is the relationship between migration and housing prices. (Descriptive, Predictive)

2. How does the effect on housing prices vary according to counties? (Descriptive)

3. The distribution patterns of migration and housing prices across states and time. (Descriptive)

4. The impact of urbanization driven by migration on housing prices. (Descriptive)



### Data Description

#### Migration Data Sets 

(https://www.irs.gov/statistics/soi-tax-stats-migration-data-2013-2020)
The Internal Revenue Service (IRS) provided the data. The migration is based on year-to-year address changes reported on individual income tax returns filed with the IRS. We use county-level data from 2013 to 2020. The migration data consists of both the inflows and outflows of people within and outside the country. 
N1 is the number of returns filed, approximating the number of migrated households. 
N2: number of personal exemptions claimed. Which approximates the number of individuals. The size of adjusted gross income (AGI) and the age of the primary taxpayer determine Agi, the aggregate migration flow at the state level. It is reported in thousands of dollars.



#### Housing Prices Data Set

(https://www.fhfa.gov/sites/default/files/2024-09/UADAggs_county_v3_2.zip)
The Federal Housing Finance Agency (FHFA) provided the housing price data set. We used the uniform appraisal dataset—aggregate statistics on appraisal—from the FHFA. An appraisal is an assessment of a property's value performed by a certified appraiser.
 It is quarterly data. The data spans from 2021- to 2022-year period.
Value is the worth of the statistical series.
Quarter – 1 = January – March
2 = April-June
3 = July-September.
4 = October-December
5 = all four quarters (yearly)
NB: For this research purpose we are using 5 which represent annual housing value.
Geoname is the formal name of a geographic unit (counties).


#### Prime Interest Rates

The control variable is interest rate from 2013 to 2020.  The interest rates were scraped from the JP Morgan Bank website and averaged by year.



### How Migration data and Housing data are related.

There is a close relationship between migration data and the housing data set. Migration patterns can directly influence the demand for housing. Areas facing elevated levels of migration may see increased development to meet demand. This can affect housing prices depending on how fast supply can catch up.

The two data sets were merged by year and the state and county FIPS code.



### Data Processing

#### Migration Data

There were no missing values. We changed the column names to make them more descriptive. We kept only county-level inflow data in the data set. Each year has its own file. Because the migration data is based on tax return filings, it may not accurately reflect the actual migration into the county, as some individuals who move in and out may not file a tax return.


#### Housing Prices Data

The Federal Housing Finance Agency "suppressed" some housing prices for counties with fewer than five values, even though we had no missing values. However, since our focus is on the housing market, we used all data from purchase prices in the data set for housing values, excluding refinance appraisals. The final data set included only data for the fifty states and the District of Columbia. We subset the housing price data by year, using the state and county FIPS codes as keys to join each year's migration data. We bound all the years into a single data frame for analysis, adding the average prime interest rate for each year as a separate column. The skewed distributions of both the migration and housing price data prompted a log transformation.
`


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = FALSE, warning = FALSE, results = 'hide')
```


```{r}

#load libraries used

library(tidyverse)
library(stringr)
library(readr)
library(readxl)
library(dplyr)
library(skimr)
library(binsreg)
library(splines)
library(sf)
library(estimatr)
library(fixest)
library(viridisLite)
library(plm)
library(broom)
library(stargazer)
library(gridExtra)
```


```{r}
# Loading the data 
# working directory needs to be set to the GitHub folder for the project
# load was not working with relative address, had to make it a full path name address

load("C:/Users/attaa/OneDrive/Desktop/r/Project-ECNS-560/Cleaned Data/clean_data_PR.RData")
clean_data = clean_data_PR  #The code below uses "clean_data" as the data frame name

```


```{r, fig.keep = "none"}
# Apply log transformation to VALUE and NO_OF_INDIVIDUALS(INFLOW) to create new log-transformed variables
clean_data$LOG_VALUE = log(clean_data$VALUE + 1)
clean_data$LOG_MIGRATION = log(clean_data$`NO_OF_INDIVIDUALS(INFLOW)` + 1)
```



#### Finding 1: The relationship between migration and housing prices

```{r}
# Create a scatter plot of LOG_MIGRATION vs. LOG_VALUE with a smoothing line
ggplot(clean_data, aes(x=LOG_MIGRATION, y =LOG_VALUE)) +
# Add points with low transparent  
  geom_point(alpha = 0.1)+
  labs(title = "Immigration and Housing Prices Relationship", x="Migration (Log Transformed)", y = "Housing Prices (Log Transformed)") +
# Add a smooth line to show trend  
  geom_smooth()
```


There is a clear positive non-linear relationship (upward-sloping pattern) between migration and housing prices. This indicates that as the number of migrants increases, housing prices also tend to increase. The points are tightly clustered around a linear trend suggesting a high correlation.


#### Finding 2:  How does the effect on housing prices vary according to counties? (Descriptive)

```{r, fig.width=10, fig.height=4}

# Plot LOG_MIGRATION over YEAR for each county, with points colored by county name
Top_Immigration = clean_data |>
  group_by(Y1_COUNTYNAME.x) |>
  summarise(mean_value = mean(`NO_OF_INDIVIDUALS(INFLOW)`, na.rm = T))|>  # Calculate the mean of migration for each county, excluding NA values
  arrange(desc(mean_value))|> # Arrange counties in descending order of mean_value
# Select the top 10 counties with the highest mean_value  
  top_n(10,mean_value)

# Filters the original data to only include top counties
filtered_counties= clean_data|>
  filter(Y1_COUNTYNAME.x %in% Top_Immigration$Y1_COUNTYNAME.x)


# Find top ten counties for house prices
top_counties_prices = clean_data |>
  group_by(Y1_COUNTYNAME.x)|># Group data by county name
  summarise(mean_value = mean(VALUE, na.rm = T))|> # # Calculate the mean of housing prices for each county, excluding NA values
  arrange(desc(mean_value))|> # Arrange counties in descending order of mean_value
# Select the top 10 counties with the highest mean_value  
  top_n(10,mean_value)

# Filters the original data to only include top counties
filtered_counties= clean_data|>
  filter(Y1_COUNTYNAME.x %in% top_counties_prices$Y1_COUNTYNAME.x) |>
  filter(STATEPOSTAL != "GA")  # Keep only Harris County, TX

# Plot Housing Prices (VALUE) over YEAR for each county, with points colored by county name
plot1 = ggplot(filtered_counties, aes(x = YEAR, y = VALUE, 
  # Define grouping by county                     
                       group = Y1_COUNTYNAME.x, 
  # Set color by county name to differentiate points by county
                       color = Y1_COUNTYNAME.x)) +
  #Add points to represent housing prices at each YEAR
  geom_point(size = 1.5) + geom_line(size = 1.2) +
  scale_color_viridis_d(option = "viridis") + theme_minimal() +
  labs(title = "Trend of Housing Prices by Year, Top 10 Counties", x = "YEAR", y = "Housing Prices", color = "County")


plot2 = ggplot(filtered_counties, aes(x = YEAR, y = `NO_OF_INDIVIDUALS(INFLOW)`,
          # Define grouping by county
                       group = Y1_COUNTYNAME.x,
  # Set color by county name to differentiate points by county
                       color = Y1_COUNTYNAME.x)) +
  #Add points to represent migration at each YEAR
  geom_point(size = 1.5) + geom_line(size = 1.2) +
  scale_color_viridis_d(option = "viridis") + theme_minimal() +
  labs(title = "Top 10 Counties for Migration by Year, Top 10 Counties", x = "YEAR", y = "Migration", color = "County")

grid.arrange(plot1, plot2, ncol = 2)
  
```


According to the house price graph, Riverside County experienced the most significant and consistent increase in housing prices, followed by Los Angeles County, which also sees a steady upward trend. King County and Harris County show moderate increases over the years. Other counties, such as San Bernardino, Tarrant, and Wake, have slower but steady growth in housing prices. Overall, there’s a clear upward trend in housing prices across most counties, with notable differences in the rate of increase.


Most of the counties with the highest migration inflow vary by year in similar ways. Los Angeles County consistently has the highest migration inflow, although it declines over time. New York County also sees a significant decrease in inflow. Some counties, like Maricopa and Riverside, show gradual increases or stable trends in migration inflow toward 2020. There is notable fluctuation in migration patterns across counties during the observed years. For example, in 2015, most of the counties experienced lower migration, but in 2016–2017, most of the top ten counties saw increases in annual immigration. This implies that a factor associated with each year could influence immigration in all these counties.




#### Finding 3: The distribution patterns of migration and housing prices across states and time. (Descriptive)

```{r}

# By the State level
# Calculates mean LOG_VALUE(Housing prices) and mean LOG_MIGRATION(Migration) for each state
state_summary2 = clean_data |>
# Group data by state postal cod  
  group_by(STATEPOSTAL) |>
  summarise(
    mean_value = mean(LOG_VALUE, na.rm = TRUE),# Calculate mean of LOG_VALUE for each state
    mean_inflow = mean(LOG_MIGRATION, na.rm = TRUE) # Calculate mean of LOG_MIGRATION for each state
  )
```


```{r}

#Taking the top 10 states for both Value and number of individuals inflow

# Filter for the top 10 states by mean VALUE and mean inflow
top_value_states = state_summary2 |>
# Arrange states in descending order of mean_value  
  arrange(desc(mean_value)) |>
 # Select the first 10 rows (top 10 states)  
  slice(1:10)

top_inflow_states = state_summary2 |>
# Arrange states in descending order of mean_inflow  
  arrange(desc(mean_inflow)) |>
 # Select the first 10 rows (top 10 states)  
  slice(1:10)

```



```{r, fig.width=10, fig.height=4, echo=FALSE}

# Filters the original data to only include top counties
filtered_inflow_states= clean_data|>
  filter(STATEPOSTAL %in% top_inflow_states$STATEPOSTAL) |>
  group_by(STATEPOSTAL, YEAR) |>
  summarise(mean_inflow_year = mean(`NO_OF_INDIVIDUALS(INFLOW)`))

# unique(filtered_inflow_states$STATEPOSTAL)


# Filters the original data to only include top counties
filtered_value_states= clean_data|>
  filter(STATEPOSTAL %in% top_value_states$STATEPOSTAL) |>
  group_by(STATEPOSTAL, YEAR) |>
  summarise(mean_value_year = mean(VALUE))

# unique(filtered_inflow_states$STATEPOSTAL)

plot5 = ggplot(filtered_value_states, aes(x = YEAR, y = mean_value_year, group = STATEPOSTAL, color = STATEPOSTAL)) +
  # Set color by county name to differentiate points by county
                       
  #Add points to represent LOG_MIGRATION at each YEAR
  geom_point(size = 1.5) + geom_line(size = 1.2) +
  scale_color_viridis_d(option = "viridis") + theme_minimal() +
  labs(title = "Top 10 States for Housing Prices by Year", x = "YEAR", y = "Housing Prices", color = "State")


plot6 = ggplot(filtered_inflow_states, aes(x = YEAR, y = mean_inflow_year, group = STATEPOSTAL, color = STATEPOSTAL)) +
  # Set color by county name to differentiate points by county
                       
  #Add points to represent LOG_MIGRATION at each YEAR
  geom_point(size = 1.5) + geom_line(size = 1.2) +
  scale_color_viridis_d(option = "viridis") + theme_minimal() +
  labs(title = "Top 10 States for Migration by Year", x = "YEAR", y = "Migration", color = "State")

grid.arrange(plot5,plot6, ncol=2)

```


From the graph above, States like Arizona (AZ) and Florida (FL) show consistently high migration levels, peaking around 2017 before declining. Migration trends might be influenced by factors like affordability, job opportunities, and quality of life.

The graph illustrates housing price trends. While prices generally increase across all states, some states, Arizona (AZ) has the highest and most consistent rise in housing prices, peaking in 2018, followed by California (CA), which also sees a steady increase. Other states, like Connecticut (CT), Delaware (DE), and Rhode Island (RI), exhibit slower but noticeable growth in housing prices over the years.

As expected, in 2014, as migration fairly declined, housing prices were stable if not declining. In 2015, as it began to rise, peaking around 2017 and declining, house prices around that period also began to rise, peaking at 2018 and fairly declining.


#### Finding 4:  Impact of urbanization on migration and housing prices

```{r,fig.width=10, fig.height=4}
# Find the top 10 counties with the highest mean of housing prices

top_counties = clean_data |>
  group_by(Y1_COUNTYNAME.x)|># Group data by county name
  summarise(mean_value = mean(LOG_VALUE, na.rm = T))|> # # Calculate the mean of housing prices for each county, excluding NA values
  arrange(desc(mean_value))|> # Arrange counties in descending order of mean_value
# Select the top 10 counties with the highest mean_value  
  top_n(10,mean_value)

# Filters the original data to only include top counties
filtered_counties= clean_data|>
  filter(Y1_COUNTYNAME.x %in% top_counties$Y1_COUNTYNAME.x)

# Box plot for top counties 
plot3 = ggplot(filtered_counties, aes(x = reorder(Y1_COUNTYNAME.x, -LOG_VALUE), y = LOG_VALUE)) +
  geom_boxplot(fill = "lightgreen") +
  labs(title = "Box Plot of Housing Prices for Top 10 Counties", x = "County", y = "Housing Prices (Log Transformed)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot the top ten states based of mean of VALUE
# Find the top 10 counties with the highest mean of inflow (migration)
top2_counties = clean_data |>
  group_by(Y1_COUNTYNAME.x)|># Group data by county name
  summarise(mean_value = mean(LOG_MIGRATION, na.rm = T))|> # Calculate the mean of migration (inflow) for each county, excluding NA values
  arrange(desc(mean_value))|> # Arrange counties in descending order of mean_value
#Select the top 10 counties with the highest mean_value  
  top_n(10,mean_value)

# Filters the original data to only include top counties
filter_counties= clean_data|>
  filter(Y1_COUNTYNAME.x %in% top2_counties$Y1_COUNTYNAME.x)

# Box plot for top counties 
plot4 = ggplot(filter_counties, aes(x = reorder(Y1_COUNTYNAME.x, -LOG_MIGRATION), y = LOG_MIGRATION)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Box Plot of Migration for Top 10 Counties", x = "County", y = "Migration (Log Transformed") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

grid.arrange(plot3, plot4, ncol = 2)
```


The boxes show the interquartile range (IQR), which is the range from the 25th percentile (bottom edge of the box) to the 75th percentile (top edge of the box).The line within each box represents the median house price for that county. Whiskers extend from the box to indicate the range of house prices within 1.5 times the IQR. Individual dots represent potential outliers that fall outside this range. 

Counties like Maricopa and Los Angeles show higher median housing prices, with Maricopa having the widest range. Riverside and King Counties also have relatively high prices, but their distributions are more compact.There are some notable outliers in counties like Maricopa and San Bernardino, suggesting extreme housing price variations in these areas.

The migration box plot above provides a summary of the migration  for the top 10 counties.
Los Angeles County leads with the highest median migration levels, followed by Maricopa County and San Diego County. 
Counties like New York County and Broward County show lower median migration and a narrower range, suggesting more consistent migration numbers with fewer outliers.

The two diagrams highlight key factors in urbanization, which is brought about by the relationship between migration and housing prices. As was shown for Maricopa and Riverside Counties, rapid population inflow means a greater demand for housing and increased prices. This shows how economic opportunities are concentrated in urban areas.
On the other hand, counties such as San Bernardino, with a relatively low housing cost but high migration, are a result of the spillover effect of urbanization. These areas likely represent the suburban areas of larger cities and thus offer relatively more affordable housing without losing the benefits of proximity to the city center.




### Econometric Analysis

### Motivation

Many factors that may not be easily quantifiable influence individual migration decisions. To provide a quantitative perspective on the relationship between housing price and migration, we conducted a regression analysis. We cannot specifically claim causal inference because of a lack of clear counterfactuals,even though we controlled for some of the variations within state and year and also controlled for prime rates across years. Therefore, the obtained results do not prove causation, but rather they depict a consistent pattern within the data sets. 



### Method
To estimate the impact of migration on housing prices, we used a linear fixed effects model that includes average prime rates as a control variable to account for other variation that the state and year fixed effects could not capture.


```{r}
# Rename columns in the clean_data_PR dataset
clean_data = clean_data|>
  rename(
# Rename the column "VALUE" to "HOUSING_PRICE"    
    HOUSING_PRICE = VALUE,
# Rename the column "NO_OF_INDIVIDUALS(INFLOW)" to "MIGRATION"
    MIGRATION = `NO_OF_INDIVIDUALS(INFLOW)`
  )
```


```{r}
# Add new columns with log-transformed values to the clean_data_PR dataset
clean_data = clean_data |>
  mutate(
# Create a new column "LOG_HOUSING_PRICE" by applying the natural log to "HOUSING_PRICE"    
    LOG_HOUSING_PRICE = log(HOUSING_PRICE),
# Create a new column "LOG_MIGRATION" by applying the natural log to "MIGRATION"
    LOG_MIGRATION = log(MIGRATION)
  )

```



```{r}
# Fit a simple linear regression model
# LOG_HOUSING_PRICE is the dependent variable (outcome).
# LOG_MIGRATION is the independent variable (predictor).
simple_reg= lm(LOG_HOUSING_PRICE~ LOG_MIGRATION, data=clean_data)

# Summarize the regression model results in a tidy format
tidy(simple_reg)
```


Using the within estimator fixed effects
```{r}
# Fit a fixed-effects regression model using plm
fix_reg = plm(
  LOG_HOUSING_PRICE ~ LOG_MIGRATION, 
# Specifies the panel data structure: "STATEPOSTAL" as cross-sectional index and "YEAR" as time index  
  index = c("STATEPOSTAL", "YEAR"),
# Use the "within" model for fixed-effects estimation
  model = "within", 
# specifies the dataset containing the variables
  data=clean_data
  )

# Summarize the regression model results in a tidy format
tidy(fix_reg)
```

With control variable  
```{r}
fix_reg_control = plm(
  LOG_HOUSING_PRICE ~ LOG_MIGRATION + AvePR, 
#Defines the panel data structure (STATEPOSTAL for states, YEAR for time periods)  
  index = c("STATEPOSTAL", "YEAR"),
#Specifies a fixed-effects model  
  model = "within", 
#Specifies the dataset containing the variables for the analysis
  data=clean_data
  )

# Summarize the regression results in a tidy format
tidy(fix_reg_control)
```

### Results

```{r, results='asis'}
# Generate a formatted regression table
stargazer(
#The regression model object from plm  
  fix_reg_control,
#Specifies the output format  
  type = "html",
#Title for the regression table
  title = "Regression Output",
#Custom label for the dependent variable in the output table
  ep.var.labels = "Log of Housing Price",
#Custom labels for the independent variables
  covariate.labels = c("Log of Migration", "Average Prime Rate"),
  digits = 3
)
```


### Interpretation
From the results, we can tell areas with higher levels of migration also have higher housing prices. As migration increases by 1 percent, housing prices also increase by approximately 1.14 percent, all other factors held constant. The data reveals the crucial correlation between migration and housing prices. The P value associated with this estimate is <0.01, which is below the conventional threshold for significance of 0.005. This means that the estimate is significant and reliable. 

Average prime rate: A unit increase in the average prime rate, is associated with approximately a 22 percent increase in housing prices, cetaris paribus. The p-value associated with this estimate is <0.01, which is highly significant, indicating that average rates and housing prices are highly correlated.


### Limitations

Our data analysis encounters several limitations that could potentially undermine the results of our regression. First, omitted variable bias. Other variables, not captured in our data set and subsequent regression analysis, may influence the effect of housing price increases. Variables such as income, employment rates, etc. may significantly influence housing prices. Also, the analysis shows relationships between migration, prime rates, and housing prices, but it does not prove that one causes the other. Other unmeasured factors, like government policies, economic shocks, or local market conditions, could influence these variables. Lastly, we adopted the fixed effects model in estimating our analysis. While fixed effects control for unchanging factors within each state and year, they do not account for time-varying factors unique to individual states, which might still influence housing prices.


### Conclusion

The influx of individuals into an area typically increases the demand for housing and often leads to higher prices due to limited supply and competition. Economic growth also correlates with migration, as people gravitate towards regions offering better job opportunities and rising wages, thereby intensifying the demand for housing. In some cases, migration can instigate gentrification, where wealthier newcomers raise property values in lower-income neighborhoods, potentially displacing long-term residents. Higher average prime rates often occur during periods of economic growth and inflation, which can push housing prices up. In a growing economy, despite rising costs of borrowing, buyers may still be willing to pay higher prices because their incomes and investments are also increasing. Sometimes, rising prime rates signal tighter monetary policies. This can lead to a temporary spike in housing prices as buyers rush to secure loans before rates climb even higher.


### Policy Implications

To address rising housing prices driven by migration, local governments should focus on building more affordable housing, rezoning for higher-density areas, and encouraging people to move to less crowded regions by improving infrastructure and job opportunities. To manage the effects of rising prime rates, governments can provide low-interest loans or subsidies for first-time buyers to keep housing affordable. Additionally, stricter lending standards can help prevent risky borrowing and speculative bubbles that drive prices even higher.


ChatGPT - For troubleshooting




