---
title: "Project Report"
author: "Simon Antwi and Steven Durtka"
date: "2024-12-5"
output: html_document
---

### Topic

### The impact of Migration on Housing Prices in the US

### Motivation

Migration is the movement of people from one place to another. This can happen internationally when people move from one country to another, or it can happen internally as people move within the same country.Migration can be temporary or permanent, voluntary or forced.It can also be seasonal.

Numerous factors, such as economic opportunities, escaping conflict or persecution, reuniting with family, or educational purposes, inform migration decisions for people moving between nations. Family, employment, educational pursuits, or other interests are some of the underlying reasons people may choose to relocate within the U.S. 

Housing prices stand for the cost of purchasing properties, such as houses, apartments, or condominiums. Immigration significantly impacts all facets of the US economy, including the housing market. We want to investigate the impact of immigrants’ influx (both domestic and foreign) on local housing prices, considering also the demand and supply factors in the market such as interest rates and general population increases.



### Why It Matters

Firstly, it is crucial for policymakers as it aids them in making informed decisions regarding immigration, which in turn helps to stabilize housing prices, ensuring affordability and a high standard of living. 

Furthermore, it can facilitate the establishment of a stable market, as both homeowners and renters have a reasonable insight into the fluctuations in the housing market. 

Finally, it can also function as a guide for urban planning and infrastructure development.



### Research Questions

1. What is the relationship between migration and housing prices. (Descriptive, Predictive)

2. How does the effect on housing prices vary according to counties? (Descriptive)

3. The distribution patterns of migration and housing prices across states and time. (Descriptive)

4. The impact of urbanization driven by migration on housing prices. (Descriptive)



### Data Description

#### Migration Data Sets 

(https://www.irs.gov/statistics/soi-tax-stats-migration-data-2013-2020)
The Internal Revenue Service (IRS) provided the data. The migration is based on year-to-year address changes reported on individual income tax returns filed with the IRS. We use county-level data from 2013 to 2020. The migration data consists of both the inflows and outflows of people within and outside the country. 
N1 is the number of returns filed, approximating the number of migrated households. 
N2: number of personal exemptions claimed. Which approximates the number of individuals. The size of adjusted gross income (AGI) and the age of the primary taxpayer determine Agi, the aggregate migration flow at the state level. It is reported in thousands of dollars.

#### Housing Prices Data Set

(https://www.fhfa.gov/sites/default/files/2024-09/UADAggs_county_v3_2.zip)
The Federal Housing Finance Agency (FHFA) provided the housing price data set. We used the uniform appraisal dataset—aggregate statistics on appraisal—from the FHFA. An appraisal is an assessment of a property's value performed by a certified appraiser.
 It is quarterly data. The data spans from 2021- to 2022-year period.
Value is the worth of the statistical series.
Quarter – 1 = January – March
2 = April-June
3 = July-September.
4 = October-December
5 = all four quarters (yearly)
NB: For this research purpose we are using 5 which represent annual housing value.
Geoname is the formal name of a geographic unit (counties).

#### Prime Interest Rates

The control variable is interest rate from 2013 to 2020.  The interest rates were scraped from the JP Morgan Bank website and averaged by year.

### How Migration data and Housing data are related.

There is a close relationship between migration data and the housing data set. Migration patterns can directly influence the demand for housing. Areas facing elevated levels of immigration may see increased development to meet demand. This can affect housing prices depending on how fast supply can catch up.

The two data sets were merged by year and the state and county FIPS code.

### Data Processing

#### Migration Data

There were no missing values. We changed the column names to make them more descriptive. We kept only county-level inflow data in the data set. Each year has its own file. Because the migration data is based on tax return filings, it may not accurately reflect the actual migration into the county, as some individuals who move in and out may not file a tax return.

#### Housing Prices Data

The Federal Housing Finance Agency "suppressed" some housing prices for counties with fewer than five values, even though we had no missing values. However, since our focus is on the housing market, we used all data from purchase prices in the data set for housing values, excluding refinance appraisals. The final data set included only data for the fifty states and the District of Columbia. We subset the housing price data by year, using the state and county FIPS codes as keys to join each year's migration data. We bound all the years into a single data frame for analysis, adding the average prime interest rate for each year as a separate column. The skewed distributions of both the migration and housing price data prompted a log transformation.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = 'hide')
```


```{r}

#load libraries used

library(tidyverse)
library(stringr)
library(readr)
library(readxl)
library(dplyr)
library(skimr)
library(binsreg)
library(splines)
library(sf)
library(estimatr)
library(fixest)
library(viridisLite)
library(gridExtra)
```


```{r}
# Loading the data 
# working directory needs to be set to the GitHub folder for the project
# load was not working with relative address, had to make it a full path name address

load("C:/Users/attaa/OneDrive/Desktop/r/Project-ECNS-560/Cleaned Data/clean_data_PR.RData")
clean_data = clean_data_PR  #The code below uses "clean_data" as the data frame name

```


```{r}
#statistical summary of the data set
summary(clean_data)
```


```{r}
skim(clean_data)

```

### Exploratory Analysis

```{r, fig.keep = "none"}
# Create a histogram of the 'VALUE' column in the 'clean_data' data 
ggplot(clean_data,aes(x=VALUE))+ 
  geom_histogram(boundary=0,binwidth = 1000)

# This shows the distribution of value(housing prices) variable, which is highly right-skewed. Most of the data are concentrated on the left side of the distribution. This points out that significant portion of the values in the dataset are small in value with few proportion been large.
```


```{r, fig.keep = "none"}
# Create a histogram of the 'NO_OF_INDIVIDUALS(INFLOW)' column in the 'clean_data' data frame
ggplot(clean_data,aes(x=`NO_OF_INDIVIDUALS(INFLOW)`))+ 
  geom_histogram(boundary=0,binwidth = 2500)

#This shows the distribution of number of individual(inflow) i.e. migration variable, which highly right-skewed. Most of the data are concentrated on the left side of the distribution. 
```


```{r, fig.keep = "none"}
# Apply log transformation to VALUE and NO_OF_INDIVIDUALS(INFLOW) to create new log-transformed variables
clean_data$LOG_VALUE = log(clean_data$VALUE + 1)
clean_data$LOG_MIGRATION = log(clean_data$`NO_OF_INDIVIDUALS(INFLOW)` + 1)
```


```{r, fig.keep = "none"}
#Creating the histogram graphs and other exploratory analysis using the log variables that we have created through transformation
# Create a histogram of the 'LOG_VALUE' column in the 'clean_data' data frame
ggplot(clean_data,aes(x=LOG_VALUE))+ 
  geom_histogram(boundary=0,binwidth = 0.5)
```

```{r, fig.keep = "none"}
# Create a histogram of the 'LOG_MIGRATION' column in the 'clean_data' data frame

ggplot(clean_data,aes(x=LOG_MIGRATION))+ 
  geom_histogram(boundary=0,binwidth = 0.5)
```





#### Finding 1: The relationship between migration and housing prices

```{r}
# Create a scatter plot of LOG_MIGRATION vs. LOG_VALUE with a smoothing line
ggplot(clean_data, aes(x=LOG_MIGRATION, y =LOG_VALUE)) +
# Add points with low transparent  
  geom_point(alpha = 0.1)+
  labs(title = "Immigration and Housing Prices Relationship", x="Migration (Log Transformed)", y = "Housing Prices (Log Transformed)") +
# Add a smooth line to show trend  
  geom_smooth()
```

There is a clear positive non-linear relationship (upward-sloping pattern) between migration and housing prices. This indicates that as the number of migrants increases, housing prices also tend to increase. The points are tightly clustered around a linear trend suggesting a high correlation.


```{r}

# However, the plot shows that the relationship is not perfectly linear.  To attempt to capture the curves in the graph, a cubic spline with 5 degrees of freedom was conducted.

#Cubic Splines
# Perform a piece wise cubic regression using B-splines
# pw_cubic=feols(VALUE~ splines::bs(`NO_OF_INDIVIDUALS(INFLOW)`,df=5),
#                clean_data)
# # Display the summary of the regression model
# summary(pw_cubic)
# 
# 
# # Generate fitted values for cubic splines using predict
# pw_cubic_predict = predict(pw_cubic,data=clean_data, interval = "confidence")
# pw_cubic_fitted=cbind(clean_data,pw_cubic_predict)
# 
# # Create a plot to visualize the piece wise cubic regression fit and confidence intervals
# ggplot(pw_cubic_fitted, aes(x = `NO_OF_INDIVIDUALS(INFLOW)`)) + theme_light()+
#   # Scatter plot of original data points 
#   geom_point(aes(y=VALUE),alpha = 0.1)+ 
#   # Confidence interval ribbon around the fitted line
#     geom_ribbon(aes(ymin = ci_low, ymax = ci_high), fill = "coral", alpha=2)+
#   # Fitted line from the piece wise cubic regression model
#   geom_line(aes(y =fit), size = 1,color = "green")+ 
#   labs(x = "Number of Immigrants", y = "Housing Prices", title = "Cubic Splines Regression (5 Degrees of Freedom)") +
#   theme_minimal()
#This plot shows the relationship between the number of immigrants and housing prices, with a fitted regression line using cubic splines with 5 degrees of freedom.
#The spline curve shows a non-linear positive relationship between the number of immigrants and housing prices. At lower levels of immigration, housing prices rise gradually. As the number of immigrants increases, the slope of the curve becomes steeper, indicating a faster rate of increase in housing prices.This could suggest that states with higher numbers of immigrants experience higher demand for housing, leading to increased housing prices

```





```{r, fig.keep = "none"}
#Generate a binned scatter plot of log-transformed VALUE and log-transformed NO_OF_INDIVIDUALS(INFLOW)
#binsreg(log(clean_data$VALUE),log(clean_data$`NO_OF_INDIVIDUALS(INFLOW)`))

# There is an overall positive relationship between immigration and housing prices, as indicated by the upward trend. As the NO_OF_INDIVIDUALS(INFLOW) increases, VALUE also tends to increase, suggesting a direct association(correlation) between these two variables. This also affirms the relationship between the variables by the scatter graph even though this show a non - linear relationship
```


 
```{r, fig.keep = "none"}

# Create a scatter plot with a smooth trend line using ggplot with clean_data data set, mapping 'NO_OF_INDIVIDUAL(INFLOW)' to x-axis and 'VALUE' to y-axis
ggplot(clean_data,aes(x=LOG_MIGRATION,y= LOG_VALUE)) +
  geom_point(alpha=0.3, size=1.5)+
# Add a smooth line using LOESS smoothing with linear fitting  
  geom_smooth(method = "loess")

# The graph shows a positive relationship (upward trend) between number of individuals inflow (migration) and housing prices, signifying that when migration(inflow) increases housing prices.The relationship is not a linear one but due to the smoothness some of the variations are been averaged out hence we do not have enough information about some key variations.
```


```{r, fig.keep = "none"}
# creating a bin scatter
#  Create 50 bins based on the 'NO_OF_INDIVIDUALS(INFLOW)' variable and calculate binned means
clean_data =clean_data|>
  mutate(INDIVIDUAL_bins=factor(ntile(LOG_MIGRATION, 50)))

# Group the data by the created bins and calculate the average 'NO_OF_INDIVIDUALS' and 'VALUE' within each bin
binned_data = clean_data |>
  group_by(INDIVIDUAL_bins) |>
  # Calculate the mean 'VALUE' for each bin, ignoring missing values
  mutate(VALUE_binned = mean(LOG_VALUE,na.rm = T),
# Calculate the mean 'NO_OF_INDIVIDUALS' for each bin, ignoring missing values
         INDIVIDUALS_binned = mean(LOG_MIGRATION, na.rm = T))

 
```


```{r, fig.keep = "none"}
# Creating the graph
# Create a scatter plot with the clean_data and binned data using 
ggplot
ggplot(binned_data) +
  geom_point(aes(x = LOG_MIGRATION, y = LOG_VALUE, color = INDIVIDUAL_bins), alpha = 0.3) +
  geom_point(aes(x = INDIVIDUALS_binned, y = VALUE_binned), size = 3, color = "navy") +  # Binned Averages
  labs(x = "Number of People inflow", y = "Average House Prices") +
  theme_minimal()

# The bin scatter plot suggests a positive correlation between the number of people inflow and average house prices. As the number of people inflow increases, average house prices generally tend to increase as well. This pattern suggests that higher inflows of people are associated with higher house prices, possibly due to demand pressures in areas with higher population inflow.
```
 


```{r, fig.keep = "none"}
# creating a local regression using the bin
# Using a local regression
# Create a smooth trend line plot using ggplot2
ggplot(binned_data,aes(x=LOG_MIGRATION, y= LOG_VALUE))+
  geom_smooth()# Add a smooth trend line to visualize the overall relationship between NO_OF_INDIVIDUALS and VALUE

# The CEF shows a clear linear upward sloping trend(positive relationship between migrations and housing prices). This suggests that higher levels of migrations are generally associated with higher house value.
```




#### Finding 2:  How does the effect on housing prices vary according to counties? (Descriptive)


```{r, fig.width=10, fig.height=4}

# Plot LOG_MIGRATION over YEAR for each county, with points colored by county name
Top_Immigration = clean_data |>
  group_by(Y1_COUNTYNAME.x) |>
  summarise(mean_value = mean(`NO_OF_INDIVIDUALS(INFLOW)`, na.rm = T))|> # # Calculate the mean of migration for each county, excluding NA values
  arrange(desc(mean_value))|> # Arrange counties in descending order of mean_value
# Select the top 10 counties with the highest mean_value  
  top_n(10,mean_value)

# Filters the original data to only include top counties
filtered_counties= clean_data|>
  filter(Y1_COUNTYNAME.x %in% Top_Immigration$Y1_COUNTYNAME.x)


# Find top ten counties for house prices
top_counties_prices = clean_data |>
  group_by(Y1_COUNTYNAME.x)|># Group data by county name
  summarise(mean_value = mean(VALUE, na.rm = T))|> # # Calculate the mean of housing prices for each county, excluding NA values
  arrange(desc(mean_value))|> # Arrange counties in descending order of mean_value
# Select the top 10 counties with the highest mean_value  
  top_n(10,mean_value)

# Filters the original data to only include top counties
filtered_counties= clean_data|>
  filter(Y1_COUNTYNAME.x %in% top_counties_prices$Y1_COUNTYNAME.x) |>
  filter(STATEPOSTAL != "GA")  # Keep only Harris County, TX

# Plot Housing Prices (VALUE) over YEAR for each county, with points colored by county name
plot1 = ggplot(filtered_counties, aes(x = YEAR, y = VALUE, 
  # Define grouping by county                     
                       group = Y1_COUNTYNAME.x, 
  # Set color by county name to differentiate points by county
                       color = Y1_COUNTYNAME.x)) +
  #Add points to represent housing prices at each YEAR
  geom_point(size = 1.5) + geom_line(size = 1.2) +
  scale_color_viridis_d(option = "viridis") + theme_minimal() +
  labs(title = "Trend of Housing Prices by Year, Top 10 Counties", x = "YEAR", y = "Housing Prices", color = "County")


plot2 = ggplot(filtered_counties, aes(x = YEAR, y = `NO_OF_INDIVIDUALS(INFLOW)`,
          # Define grouping by county
                       group = Y1_COUNTYNAME.x,
  # Set color by county name to differentiate points by county
                       color = Y1_COUNTYNAME.x)) +
  #Add points to represent migration at each YEAR
  geom_point(size = 1.5) + geom_line(size = 1.2) +
  scale_color_viridis_d(option = "viridis") + theme_minimal() +
  labs(title = "Top 10 Counties for Migration by Year, Top 10 Counties", x = "YEAR", y = "Migration", color = "County")

grid.arrange(plot1, plot2, ncol = 2)
  
```


Riverside County experiences the most significant and consistent increase in housing prices, followed by Los Angeles County, which also sees a steady upward trend. King County and Harris County show moderate increases over the years. Other counties, such as San Bernardino, Tarrant, and Wake, have slower but steady growth in housing prices. Overall, there’s a clear upward trend in housing prices across most counties, with notable differences in the rate of increase.


Most of the counties with the highest migration inflow vary by year in similar ways. Los Angeles County consistently has the highest migration inflow, although it declines over time. New York County also sees a significant decrease in inflow. Some counties, like Maricopa and Riverside, show gradual increases or stable trends in migration inflow toward 2020. There is notable fluctuation in migration patterns across counties during the observed years. For example, in 2015, most of the counties experienced lower migration, but in 2016–2017, most of the top ten counties saw increases in annual immigration. This implies that a factor associated with each year could influence immigration in all these counties.





#### Finding 3: The distribution patterns of migration and housing prices across states and time. (Descriptive)


```{r}
#### Mean Housing Prices by state
# #Group the data by state and calculates the mean of VALUE
# state_summary = clean_data|>
#   group_by(STATEPOSTAL)|>
#   summarise(mean_value = mean(LOG_VALUE,na.rm = T))
# 
# #Plot the box plot
# ggplot(state_summary, aes(x = reorder(STATEPOSTAL, -mean_value), y= mean_value))+
#    geom_point(fill = "lightblue", size = 3) +
#   labs(title = "Average Housing Prices by State", x = "State", y = "Average House Prices (Log transformed)") +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))

```



```{r}

# By the State level
# Calculates mean LOG_VALUE(Housing prices) and mean LOG_MIGRATION(Migration) for each state
state_summary2 = clean_data |>
# Group data by state postal cod  
  group_by(STATEPOSTAL) |>
  summarise(
    mean_value = mean(LOG_VALUE, na.rm = TRUE),# Calculate mean of LOG_VALUE for each state
    mean_inflow = mean(LOG_MIGRATION, na.rm = TRUE) # Calculate mean of LOG_MIGRATION for each state
  )
```


```{r}

#Taking the top 10 states for both Value and number of individuals inflow

# Filter for the top 10 states by mean VALUE and mean inflow
top_value_states = state_summary2 |>
# Arrange states in descending order of mean_value  
  arrange(desc(mean_value)) |>
 # Select the first 10 rows (top 10 states)  
  slice(1:10)

top_inflow_states = state_summary2 |>
# Arrange states in descending order of mean_inflow  
  arrange(desc(mean_inflow)) |>
 # Select the first 10 rows (top 10 states)  
  slice(1:10)

```



```{r, fig.width=10, fig.height=4, echo=FALSE}

# Filters the original data to only include top counties
filtered_inflow_states= clean_data|>
  filter(STATEPOSTAL %in% top_inflow_states$STATEPOSTAL) |>
  group_by(STATEPOSTAL, YEAR) |>
  summarise(mean_inflow_year = mean(`NO_OF_INDIVIDUALS(INFLOW)`))

# unique(filtered_inflow_states$STATEPOSTAL)


# Filters the original data to only include top counties
filtered_value_states= clean_data|>
  filter(STATEPOSTAL %in% top_value_states$STATEPOSTAL) |>
  group_by(STATEPOSTAL, YEAR) |>
  summarise(mean_value_year = mean(VALUE))

# unique(filtered_inflow_states$STATEPOSTAL)

plot5 = ggplot(filtered_value_states, aes(x = YEAR, y = mean_value_year, group = STATEPOSTAL, color = STATEPOSTAL)) +
  # Set color by county name to differentiate points by county
                       
  #Add points to represent LOG_MIGRATION at each YEAR
  geom_point(size = 1.5) + geom_line(size = 1.2) +
  scale_color_viridis_d(option = "viridis") + theme_minimal() +
  labs(title = "Top 10 States for Housing Prices by Year", x = "YEAR", y = "Housing Prices", color = "State")


plot6 = ggplot(filtered_inflow_states, aes(x = YEAR, y = mean_inflow_year, group = STATEPOSTAL, color = STATEPOSTAL)) +
  # Set color by county name to differentiate points by county
                       
  #Add points to represent LOG_MIGRATION at each YEAR
  geom_point(size = 1.5) + geom_line(size = 1.2) +
  scale_color_viridis_d(option = "viridis") + theme_minimal() +
  labs(title = "Top 10 States for Migration by Year", x = "YEAR", y = "Migration", color = "State")

grid.arrange(plot5,plot6, ncol=2)

```


From the graph above, States like Arizona (AZ) and Florida (FL) show consistently high migration levels, peaking around 2017 before declining. Migration trends might be influenced by factors like affordability, job opportunities, and quality of life.

The graph illustrates housing price trends. While prices generally increase across all states, some states, Arizona (AZ) has the highest and most consistent rise in housing prices, peaking in 2018, followed by California (CA), which also sees a steady increase. Other states, like Connecticut (CT), Delaware (DE), and Rhode Island (RI), exhibit slower but noticeable growth in housing prices over the years.

As expected, in 2014, as migration fairly declined, housing prices were stable if not declining. In 2015, as it began to rise, peaking around 2017 and declining, house prices around that period also began to rise, peaking at 2018 and fairly declining.




#### Finding 4:  Impact of urbanization on migration and housing prices

```{r,fig.width=10, fig.height=4}
# Find the top 10 counties with the highest mean of housing prices

top_counties = clean_data |>
  group_by(Y1_COUNTYNAME.x)|># Group data by county name
  summarise(mean_value = mean(LOG_VALUE, na.rm = T))|> # # Calculate the mean of housing prices for each county, excluding NA values
  arrange(desc(mean_value))|> # Arrange counties in descending order of mean_value
# Select the top 10 counties with the highest mean_value  
  top_n(10,mean_value)

# Filters the original data to only include top counties
filtered_counties= clean_data|>
  filter(Y1_COUNTYNAME.x %in% top_counties$Y1_COUNTYNAME.x)

# Box plot for top counties 
plot3 = ggplot(filtered_counties, aes(x = reorder(Y1_COUNTYNAME.x, -LOG_VALUE), y = LOG_VALUE)) +
  geom_boxplot(fill = "lightgreen") +
  labs(title = "Box Plot of Housing Prices for Top 10 Counties", x = "County", y = "Housing Prices (Log Transformed)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot the top ten states based of mean of VALUE
# Find the top 10 counties with the highest mean of inflow (migration)
top2_counties = clean_data |>
  group_by(Y1_COUNTYNAME.x)|># Group data by county name
  summarise(mean_value = mean(LOG_MIGRATION, na.rm = T))|> # Calculate the mean of migration (inflow) for each county, excluding NA values
  arrange(desc(mean_value))|> # Arrange counties in descending order of mean_value
#Select the top 10 counties with the highest mean_value  
  top_n(10,mean_value)

# Filters the original data to only include top counties
filter_counties= clean_data|>
  filter(Y1_COUNTYNAME.x %in% top2_counties$Y1_COUNTYNAME.x)

# Box plot for top counties 
plot4 = ggplot(filter_counties, aes(x = reorder(Y1_COUNTYNAME.x, -LOG_MIGRATION), y = LOG_MIGRATION)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Box Plot of Migration for Top 10 Counties", x = "County", y = "Migration (Log Transformed") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

grid.arrange(plot3, plot4, ncol = 2)
```


The boxes show the interquartile range (IQR), which is the range from the 25th percentile (bottom edge of the box) to the 75th percentile (top edge of the box).The line within each box represents the median house price for that county. Whiskers extend from the box to indicate the range of house prices within 1.5 times the IQR. Individual dots represent potential outliers that fall outside this range. 

Counties like Maricopa and Los Angeles show higher median housing prices, with Maricopa having the widest range. Riverside and King Counties also have relatively high prices, but their distributions are more compact.There are some notable outliers in counties like Maricopa and San Bernardino, suggesting extreme housing price variations in these areas.

The migration box plot above provides a summary of the migration  for the top 10 counties.
Los Angeles County leads with the highest median migration levels, followed by Maricopa County and San Diego County. 
Counties like New York County and Broward County show lower median migration and a narrower range, suggesting more consistent migration numbers with fewer outliers.

The two diagrams highlight key factors in urbanization, which is brought about by the relationship between migration and housing prices. As was shown for Maricopa and Riverside Counties, rapid population inflow means a greater demand for housing and increased prices. This shows how economic opportunities are concentrated in urban areas.
On the other hand, counties such as San Bernardino, with a relatively low housing cost but high migration, are a result of the spillover effect of urbanization. These areas likely represent the suburban areas of larger cities and thus offer relatively more affordable housing without losing the benefits of proximity to the city center.


```{r}
# Polynomial Regression 
# # Fit a 4th degree polynomial regression model of VALUE on NO_OF_INDIVIDUALS(INFLOW)
# Poly_reg = feols(VALUE ~ poly(`NO_OF_INDIVIDUALS(INFLOW)`, 4), data = clean_data)
# 
# # Display the summary of the regression model
# summary(Poly_reg)
```

```{r}
# Predicting the fitted values
# predict to get fitted values
# poly_predict = predict(Poly_reg,data = clean_data,interval = "confidence")
# Combine the original data with the predicted values and confidence intervals
# poly_fitted = cbind(clean_data,poly_predict)
# 
# 
# # Create a plot to visualize the polynomial regression fit and confidence intervals
# ggplot(poly_fitted, aes(x =`NO_OF_INDIVIDUALS(INFLOW)`))+ theme() +
#   geom_point(aes(y=VALUE),alpha = 0.1)+  # Scatter plot of actual data 
#   # Confidence interval ribbon around the fitted line
#   geom_ribbon(aes(ymin = ci_low, ymax = ci_high), fill = "coral") +
#   # Fitted line from the polynomial regression
#   geom_line(aes(y = fit),size = 1,color = "orange") + 
#   labs(x = "Number of Immigrants", y = "Housing Prices", title = "Forth-Order Polynomial Regression") +
#   theme_minimal()
# This graph shows a scatter plot of the relationship between Number of Immigrants (on the x-axis) and Housing Prices (on the y-axis), with a fourth-order polynomial regression line (in orange) fitted to the data. The graph show an overall positive trend between the two variables, where housing prices increase as the number of immigrants grows, but with considerable spread. The polynomial regression captures all the complexities that would have otherwise be smoothen out by fitting a linear regression. At the initial stages the housing prices increase as the number of immigrants rises. The curve begins to flatten at the middle part (showing slow rate of increase in housing prices) and peaks again. And at the tail end has a slight downward slope indicating that increase in immigrant no longer had an impact on housing prices.

```




```{r}

#Bin Regression
# # Perform a binned regression using the feols function
# binned_reg = feols(LOG_VALUE~as_factor(ntile(LOG_MIGRATION,10)), data = clean_data)
# # Display the summary of the regression model
# summary(binned_reg)
# 
# # Predict values using the binned regression model and add confidence intervals
# binned_reg_predict=predict(binned_reg,data=clean_data,interval = "confidence")
# # Combine the original data with the predicted values and confidence intervals
# binned_reg_fitted= cbind(clean_data,binned_reg_predict)
# 
# # Create a plot to visualize the binned regression fit and confidence intervals
# ggplot(binned_reg_fitted,aes(x=LOG_MIGRATION))+ theme_light() +
#    # Scatter plot of actual data points with low opacity
#   geom_point(aes(y = LOG_VALUE), alpha = 0.1) +
#    # Confidence interval ribbon around the fitted line
#   geom_ribbon(aes(ymin = ci_low, ymax = ci_high), fill = "red") +
#   # Fitted line from the binned regression model
#   geom_line(aes(y = fit), size = 1, color = "lightgreen")

# The scatter plot provides a clear relationship between the housing prices and migration. The dense cloud of points indicates a positive correlation between migration and housing prices. As migration increases, housing prices also tends to increase, showing a general upward trend.

```


```{r}
# Reshape data to a long format for easier plotting of multiple variables

# clean_data_long = clean_data |>
#   pivot_longer(
#     cols = c(LOG_MIGRATION, LOG_VALUE),# Select columns to pivot into long format
#     names_to = "Variable", # Name the new column that will hold the variables names
#     values_to = "Log_Value" # Name the new column that will hold the values of the variables
#     )
# 
# # Plot density distributions for LOG_MIGRATION (Migration) and LOG_VALUE (Housing prices)
# ggplot(clean_data_long, aes(x = Log_Value, fill = Variable)) +
# # Density plot with transparency and smoother curve  
#   geom_density(alpha = 0.4, adjust = 1.2) +
# #Adds plot title  
#   labs(title = "Density Plot of LOG_MIGRATION and LOG_VALUE",
#        x = "Log Values",
#        y = "Density") +
# # Apply a minimal theme for a clear visualization  
#   theme_minimal() +
# # Separate panels for each variable, with independent scales  
#   facet_wrap(~ Variable, scales = "free")  # Separate panels for each variable

# Create a variable if immigration is greater than the median
# clean_data_Rev = clean_data |>
#   mutate(High_Imm = ifelse(LOG_MIGRATION > median(LOG_MIGRATION), TRUE, FALSE))
# 
# ggplot(clean_data_Rev, aes(LOG_VALUE, fill = High_Imm, na.rm=TRUE)) + 
#   geom_density(alpha = 0.4) +
#   labs(title = "Housing Prices, by Immigration", x="Prices (Ln Transformed)", y = "Density", fill = "High Immigration")

# High immigration is defined as inflow greater than the median. The density of housing prices is generally greater for counties with immigration above the median inflow.


```



